### 2015-02-01 paper/genE6SSM/EWSB/TeX/genE6SSM-EWSB.tex
Added the SARAH generated TeX files for the RGEs

\documentclass[A4]{article} 
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{feynmp}
\usepackage{hyperref}
\usepackage{longtable}
\DeclareGraphicsRule{*}{mps}{*}{}
\graphicspath{{./Diagrams/}}
\newcommand{\Tr}{\mbox{Tr}}
\textwidth 18cm
\oddsidemargin -1.0cm
\evensidemargin -1.0cm
\renewcommand{\arraystretch}{1.3} 
\setlength{\parindent}{0em} 
\begin{document}
\title{Generic E6SSM\\ Superpotential, Rotations and Interactions for eigenstates {\tt'EWSB'} \\ including Renormalization Group Equations \\ 
including one-loop Self-Energies \\ 
} 
\author{SARAH 4.2.1} 
 \maketitle 
 \vspace{10cm} 
This file was automatically generated by SARAH version 4.2.1.  \\ 
References: {\bf arXiv: 1309.7223 }, {\bf Comput.Phys.Commun.184:1792-1809,2011 (1207.0906) }, {\bf Comput.Phys.Commun.182:808-833,2011 (1002.0840) }, {\bf Comput.Phys.Commun.181:1077-1086,2010 (0909.2863) }, {\bf arXiv: 0806.0538 } \\ 
Package Homepage: projects.hepforge.org/sarah/ \\ 
by {\bf Florian Staub, fnstaub@th.physik.uni-bonn.de} 
 \pagebreak 
 \tableofcontents 
 \pagebreak 
 \allowdisplaybreaks 
\input{SF}
\section{Renormalization Group Equations}
\input{rges} 
\section{Field Rotations}
\input{RotationsGaugeEWSB} 
\input{matricesScalarsEWSB} 
\input{matricesFermionsEWSB} 
\section{Vacuum Expectation Values}
\input{VEVsEWSB} 
\section{Tadpole Equations}
\input{TadpoleEquations} 
\section{Particle content for eigenstates {\tt 'EWSB'}}
\input{particles} 
\section{One Loop Self-Energy and One Loop Tadpoles for eigenstates {\tt 'EWSB'}}
\subsection{One Loop Self-Energy}
\input{SelfEnergy1Loop} 
\subsection{Tadpoles}
\input{Tadpoles1Loop} 
\section{Interactions for eigenstates {\tt 'EWSB'}}
\subsection{Three Scalar-Interaction}
\input{VerticesSSS}
\subsection{Two Scalar-One Vector Boson-Interaction}
\input{VerticesSSV}
\subsection{One Scalar-Two Vector Boson-Interaction}
\input{VerticesSVV}
\subsection{Two Fermion-One Vector Boson-Interaction}
\input{VerticesFFV}
\subsection{Two Fermion-One Scalar Boson-Interaction}
\input{VerticesFFS}
\subsection{Three Vector Boson-Interaction}
\input{VerticesVVV}
\subsection{Four Scalar-Interaction}
\input{VerticesSSSS}
\subsection{Two Scalar-Two Vector Boson-Interaction}
\input{VerticesSSVV}
\subsection{Four Vector Boson-Interaction}
\input{VerticesVVVV}
\subsection{Two Ghosts-One Vector Boson-Interaction}
\input{VerticesGGV}
\subsection{Two Ghosts-One Scalar-Interaction}
\input{VerticesGGS}
\input{CG}
\end{document}